Finiteness in Cubical Type Theory

### 2020-07-01 sections/finiteness-predicates.tex
finished applying edits

To actually use values of this type we have the following two eliminators:
\begin{multicols}{2} \null \vfill
  \begin{agdalisting} \label{elim-prop}
    \ExecuteMetaData[agda/HITs/PropositionalTruncation/Display.tex]{rec-prop-trunc}
  \end{agdalisting} \vfill \null \columnbreak \null \vfill
  \begin{agdalisting} \label{elim-prop-coh}
    \ExecuteMetaData[agda/HITs/PropositionalTruncation/Display.tex]{rec-prop-trunc-set}
  \end{agdalisting} \vfill \null
\end{multicols} \noindent
While these are not the only eliminators available for \proptrunc{A}, they are
the only ones we will use in this paper.
The first says we can eliminate into any proposition: interestingly, this allows
us to define a monad instance for \AgdaDatatype{\(\lVert \_ \rVert\)}, meaning
we can use things like do-notation.
The second says we can eliminate into a set as long as the function we use
doesn't care about which value it's given: formally, \(f\) in this example has
to be ``coherently constant'' \cite{krausGeneralUniversalProperty2015}.

With this, we can define cardinal finiteness:
\begin{definition}[Cardinal Finiteness]
  A type \(A\) is cardinally finite if there exists a propositionally truncated
  proof that \(A\) is manifest Bishop finite or equivalent to a finite prefix of
  the natural numbers.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinal-def}
  \end{agdalisting}
\end{definition}
\subsection{Deriving Uniquely-Determined Quantities}
At first glance, it might seem that we lose any useful properties we could
derive from \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\).
Luckily, this is not the case: we will show here how to use the eliminators from
propositional truncation (Equations \ref{elim-prop} and \ref{elim-prop-coh}) to
use to of our already-proven derivations from \AgdaFunction{\ensuremath{\mathcal{B}}}
on \AgdaFunction{\ensuremath{\mathcal{C}}}. \todo{rephrase?}

The following two lemmas are proven in
\cite{yorgeyCombinatorialSpeciesLabelled2014} (Proposition 2.4.9 and 2.4.10,
respectively), in much the same way as we have done here.
Our contribution for this section is simply the formalisation.
\begin{lemma}
  Given a cardinally finite type, we can derive the type's cardinality, as well
  as a propositionally truncated proof of equivalence with \AgdaDatatype{Fin}s of
  the same cardinality.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinality-is-unique}
  \end{agdalisting}
\end{lemma}
\begin{proof}
  The high-level overview of our proof is as follows:
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinality-is-unique-impl}
  \end{agdalisting}
  It is the composition of two operations: first, with
  \AgdaFunction{\ensuremath{\lVert \text{map} \rVert}}, we change the truncated
  proof of manifest bishop finiteness to a proof of equivalence with fin.

@@

  The tricky part of the proof is \AgdaFunction{const-alg}: here we need to show
  that \AgdaFunction{alg} returns the same value no matter its input.
  That output is a pair, the first component of which is the cardinality, and
  the second the truncated equivalence proof.
  The truncated proofs in the output are trivially equal by the truncation, so
  our obligation now has been reduced to:
  \begin{equation}
    \frac{(n :\;\AgdaDatatype{\ensuremath{\mathbb{N}}}) \; \; \; (p : \AgdaDatatype{Fin}\;n\;\AgdaFunction{\ensuremath{\simeq}}\;A) \; \; \;
      (m :\;\AgdaDatatype{\ensuremath{\mathbb{N}}}) \; \; \; (q : \AgdaDatatype{Fin}\;m\;\AgdaFunction{\ensuremath{\simeq}}\;A)
    }{
      n\;\AgdaFunction{\ensuremath{\equiv}}\;m
    }
  \end{equation}
  Given univalence we have \(\AgdaDatatype{Fin}\;n \;\AgdaFunction{\ensuremath{\equiv}}\; \AgdaDatatype{Fin}\;m\),
  and the rest of our task is to prove:
  \begin{equation}
    \frac{\AgdaDatatype{Fin}\;n\;\AgdaDatatype{\ensuremath{\equiv}}\;\AgdaDatatype{Fin}\;m}{n \equiv m}
  \end{equation}

  This is a well-known chestnut in dependently-typed programming, and one that
  has a surprisingly tricky and complex proof.
  We do not include it here, since it has already been explored elsewhere, but
  it is present in our formalisation.
\end{proof}

Another proof that carries over from manifest Bishop finiteness is decidable
equality.
This demonstrates an interesting aspect of propositional truncation: it can
contain computational information.
That is in contrast to other ways to ``truncate'' types: \(\neg \neg P\), for
instance, is a way to provide a ``proof'' of \(P\) without revealing anything
about \(P\) in MLTT.
No matter how much we prove that a function from \(P\) doesn't care about which
\(P\) it got, though, we can never extract any kind of algorithm or computation
from \(\neg \neg P\). \todo{Is this true? It certainly seems true...}

Here we'll briefly show how to extract ``computation'' from under a
propositional truncation.
\begin{lemma} \label{cardinal-finite-discrete}
  Any cardinal-finite set has decidable equality.
\end{lemma}
\begin{proof}
We can extract plenty of computation from under propositional truncation as long
as we prove the thing we're extracting is a proposition (Equation~\ref{elim-prop}). 
Our task, in other words, is to prove the following:
\begin{agdalisting}
  \ExecuteMetaData[agda/Relation/Nullary/Discrete/Properties.tex]{is-prop-discrete}
\end{agdalisting}

\AgdaInductiveConstructor{yes} is a little more interesting: its payload is
\agdaequiv{x}{y}.
How can we prove that the path between \(x\) and \(y\) is a proposition?
It turns out that there is a class of types for which all paths are
propositions: the \emph{sets}.
\begin{agdalisting} \label{isSet}
  \ExecuteMetaData[agda/Snippets/Introduction.tex]{isSet}
\end{agdalisting}
This is the next homotopy level up from the propositions (Equation~\ref{isProp}).
More importantly, there is an important theorem relating to sets which
\emph{also} relates to decidable equality: Hedberg's theorem
\cite{hedbergCoherenceTheoremMartinLof1998}.
\begin{theorem}[Hedberg's Theorem]
  Any type with decidable equality is a set.
\end{theorem} \noindent
And of course we know that \(A\) here has decidable equality: we were just given
two proofs of that fact at the beginning of this proof!
\end{proof}
\subsection{Restrictiveness}
So far our explorations into finiteness predicates have pushed us in the
direction of ``less informative'': however, as mentioned in the introduction, we
can \emph{also} ask how \emph{restrictive} certain predicates are.
Since split enumerability and manifest Bishop finiteness imply each other we
know that there can be no type which satisfies one but not the other.
We also know that manifest Bishop finiteness implies cardinal finiteness, but we
do \emph{not} have a function in the other direction:
\begin{equation}
  \agdacal{C}\;A\rightarrow\agdacal{B}\;A
\end{equation}
So the question arises naturally: is there a cardinally finite type which is
\emph{not} manifest Bishop finite?

It turns out the answer is no!
We can prove the following statement:
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{no-gap-card-bishop}
\end{agdalisting}
We use the fact that \agdabot\; is itself a proposition to eliminate from under
the truncation.
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{no-gap-card-bishop-proof}
\end{agdalisting}

This tells us that manifest bishop finiteness, cardinal finiteness, and split
enumerability all refer to the same class of types.
\subsection{Going from Cardinal Finiteness to Manifest Bishop Finiteness}
We know of course that we can convert any proof of manifest Bishop finiteness to
a proof of Cardinal finiteness: it's just the truncation function
\AgdaInductiveConstructor{\(\lvert \_ \rvert\)}.
It's the other direction which presents a difficulty:
\begin{theorem} \label{cardinal-to-manifest-bishop}
  Any cardinal finite type with a total order is Bishop finite.
\end{theorem}
The proof for this particular theorem is quite involved in the formalisation, so
we only give its sketch here.

Our strategy will be to \emph{sort} the support list of the proof for Bishop
finiteness, and then prove that the sorting function is coherently constant,
thereby satisfying the eliminator in Equation~\ref{elim-prop-coh}.
We need to show, in other words, that sorting two support lists from proofs of
manifest Bishop finiteness on the same type with the same order always returns
the same result.
The sort we use is a standard insertion sort:
\begin{multicols}{2}
  \begin{agdalisting}
    \ExecuteMetaData[agda/Data/List/Sort.tex]{insert}
  \end{agdalisting} \columnbreak \null \vfill \null \vfill
  \begin{agdalisting}
    \ExecuteMetaData[agda/Data/List/Sort.tex]{insertion-sort}
  \end{agdalisting} \vfill \null
\end{multicols} \vspace{-\baselineskip} \noindent
And we prove that \AgdaFunction{sort} produces a list which is sorted, and a
permutation of its input.
\begin{multicols}{2}
  \begin{agdalisting}
    \ExecuteMetaData[agda/Data/List/Sort.tex]{sort-sorts}
  \end{agdalisting} \columnbreak
  \begin{agdalisting}
    \ExecuteMetaData[agda/Data/List/Sort.tex]{sort-perm}
  \end{agdalisting}
\end{multicols} \vspace{-\baselineskip}
We've introduced permutations here: we take their definition from
\cite{danielssonBagEquivalenceProofRelevant2012}.
Two lists are permutations of each other if their membership proofs are all
equivalent.
\begin{agdalisting}
  \ExecuteMetaData[agda/Data/List/Relation/Binary/Permutation.tex]{perm-def}
\end{agdalisting}
This definition fits particularly well for two reasons: first, it is defined on
containers generically, which fits well with our finiteness predicates.
Secondly, it is extremely straightforward to show that the support lists of any
two proofs of manifest Bishop finiteness must be permutations of each other:
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{perm-bish}
\end{agdalisting}

Almost all of the pieces are in place now: we know that the support lists of all
proofs of \(\agdacal{B}\;A\) are permutations of each other, and we know that
\AgdaFunction{sort} returns a sorted permutation of its input.
The final piece of the puzzle is the following:
\begin{agdalisting}
  \ExecuteMetaData[agda/Data/List/Sort.tex]{sorted-perm-eq}
\end{agdalisting}
If two sorted lists are both permutations of each other they must be equal.
Connecting up all the pieces we get the following:
\begin{agdalisting}
  \ExecuteMetaData[agda/Data/List/Sort.tex]{perm-invar}
\end{agdalisting}
Because we know that all support lists of \(\agdacal{B}\;A\) are permutations of
each other this is enough to prove that \AgdaFunction{sort} is coherently
constant, and therefore can eliminate from within a truncation.
The second component of the output pair (the cover proof) follows quite
naturally from the definition of permutations.
\section{Manifest Enumerability} \label{manifest-enumerability}
We have now explored quite far in the direction of restricting information:
indeed, cardinal finiteness, a mere proposition, has the minimal amount of
information a predicate can have.
What we haven't seen, however, is a predicate with a different domain to the
others.
Since manifest Bishop finiteness and split enumerability have equal strength
\todo{ref here}, and since we've shown that there is no type which is cardinal
finite but not manifest Bishop finite, we haven't explored any of the
``restrictiveness'' axis from Figure~\ref{finite-classification}.

What we're looking for is a type which is in some sense finite, but doesn't
conform to any of the predicates we've seen so far.
The \emph{circle} is such a type:
\begin{agdalisting}\label{circle-def}
  \ExecuteMetaData[agda/Snippets/Circle.tex]{circle-def}
\end{agdalisting}
The thing that this type has which precludes it from being, say, split
enumerable, is its \emph{higher homotopy structure}. \todo{is this the right phrase?}

@@

But the type certainly \emph{seems} finite!
In order to explore the ``restrictiveness'' axis in
Figure~\ref{finite-classification}, then, we'll need to construct a predicate
which admits the circle.
Manifest enumerability is one such predicate.
\begin{definition}[Manifest Enumerability]
  Manifest enumerability is an enumeration predicate like Bishop finiteness or
  split enumerability with the only difference being a propositionally truncated
  membership proof.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable/Container.tex]{manifest-enum-def}
  \end{agdalisting}
\end{definition}

@@

And indeed this means that the circle is manifestly enumerable. \todo{This uses
  a bunch of cubical primitives!}
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{circle-is-manifest-enum}
\end{agdalisting}
\subsection{Surjections}
We already saw that split enumerability was the listed form of a split
surjection: what we didn't explain was why the word ``split'' was placed before
surjection.
In the presence of higher homotopies than sets, split surjections are actually
\emph{not} a satisfactory definition of surjection.
And we are most certainly in the presence of higher homotopies: just moments ago
we were introduced to the circle.
In these cases, the following definition of surjections is
preferred~\cite[definition 4.6.1]{hottbook}:
\begin{multicols}{2}
  \begin{agdalisting} \label{surj-eqn}
    \ExecuteMetaData[agda/Function/Surjective/Base.tex]{surjective}
  \end{agdalisting} \columnbreak
  \begin{agdalisting} \label{surj-arrow-eqn}
    \ExecuteMetaData[agda/Function/Surjective/Base.tex]{surjection}
  \end{agdalisting}
\end{multicols}

>>>>

To actually use values of this type we have the following eliminator:
\begin{agdalisting} \label{elim-prop}
  \ExecuteMetaData[agda/HITs/PropositionalTruncation/Display.tex]{rec-prop-trunc}
\end{agdalisting}
This says that we can eliminate into any proposition: interestingly, this allows
us to define a monad instance for \AgdaDatatype{\(\lVert \_ \rVert\)}, meaning
we can use things like do-notation. \todo{Bit more on do-notation etc?}

With this, we can define cardinal finiteness:
\begin{definition}[Cardinal Finiteness]
  A type \(A\) is cardinally finite if there exists a propositionally truncated
  proof that \(A\) is manifest Bishop finite or equivalent to a finite prefix of
  the natural numbers.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinal-def}
  \end{agdalisting}
\end{definition}
\subsection{Deriving Uniquely-Determined Quantities}
At first glance, it might seem that we lose any useful properties we could
derive from \(\AgdaDatatype{\ensuremath{\mathcal{B}}}\).
Luckily, this is not the case: we will show here how to derive decidable
equality (Lemma~\ref{cardinal-finite-discrete}) and cardinality
(Lemma~\ref{card-finite-cardinality}) out from under the
truncation.\todo{Rephrase?}
Those two lemmas are proven in
\cite{yorgeyCombinatorialSpeciesLabelled2014} (Proposition 2.4.9 and 2.4.10,
respectively), in much the same way as we have done here.
Our contribution for this section is simply the formalisation.

First we'll show that decidable equality carries over from manifest Bishop
finiteness.
Before we do, note that the fact that we can do this says something interesting
about propositional truncation: it has computational, or algorithmic, content.
That is in contrast to other ways to ``truncate'' types: \(\neg \neg P\), for
instance, is a way to provide a ``proof'' of \(P\) without revealing anything
about \(P\) in MLTT.
No matter how much we prove that a function from \(P\) doesn't care about which
\(P\) it got, though, we can never extract any kind of algorithm or computation
from \(\neg \neg P\). \todo{Is this true? It certainly seems true...}
\begin{lemma} \label{cardinal-finite-discrete}
  Any cardinal-finite set has decidable equality.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{card-discrete}
  \end{agdalisting}
\end{lemma}
\begin{proof}
We already know that manifest Bishop finiteness implies decidable equality
\todo{ref here}; to apply that proof to cardinal finiteness we'll use the
eliminator in Equation~\ref{elim-prop}.
Our task, in other words, is to prove the following:
\begin{agdalisting}
  \ExecuteMetaData[agda/Relation/Nullary/Discrete/Properties.tex]{is-prop-discrete}
\end{agdalisting}

\AgdaInductiveConstructor{yes} is a little more interesting: its payload is
\agdaequiv{x}{y}.
How can we prove that the path between \(x\) and \(y\) is a proposition?
It turns out that there is a class of types for which all paths are
propositions: the \emph{sets}.
\begin{agdalisting} \label{isSet}
  \ExecuteMetaData[agda/Snippets/Introduction.tex]{isSet}
\end{agdalisting}
This is the next homotopy level up from the propositions (Equation~\ref{isProp}).
More importantly, there is an important theorem relating to sets which
\emph{also} relates to decidable equality: Hedberg's theorem
\cite{hedbergCoherenceTheoremMartinLof1998}.
This tells us that any type with decidable equality is a set.
\begin{agdalisting}
  \ExecuteMetaData[agda/Relation/Nullary/Discrete/Properties.tex]{discrete-isset}
\end{agdalisting}
And of course we know that \(A\) here has decidable equality: we were just given
two proofs of that fact at the beginning of this proof!

This suffices to prove that decidable equality is itself a proposition, and
therefore that we can apply Equation~\ref{elim-prop} and the proof that bishop
finiteness implies decidable equality to cardinal finiteness, proving our goal.
\end{proof}

The next thing we can derive from underneath the truncation in cardinal
finiteness is a natural number representing the actual cardinality of the finite
type.
Of course \agdambb{N} isn't a proposition, so the eliminator in
equation~\ref{elim-prop} won't work for us here.
Instead we will use the following:
\begin{agdalisting} \label{elim-prop-coh}
  \ExecuteMetaData[agda/HITs/PropositionalTruncation/Display.tex]{rec-prop-trunc-set}
\end{agdalisting}
This says that we can eliminate into a set as long as the function we use
doesn't care about which value it's given: formally, \(f\) in this example has
to be ``coherently constant'' \cite{krausGeneralUniversalProperty2015}.

With that, we can move on to the proof:
\begin{lemma} \label{card-finite-cardinality}
  Given a cardinally finite type, we can derive the type's cardinality, as well
  as a propositionally truncated proof of equivalence with \AgdaDatatype{Fin}s of
  the same cardinality.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinality-is-unique}
  \end{agdalisting}
\end{lemma}
\begin{proof}
  The high-level overview of our proof is as follows:
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{cardinality-is-unique-impl}
  \end{agdalisting}
  It is the composition of two operations: first, with
  \AgdaFunction{\ensuremath{\lVert \text{map} \rVert}}, we change the truncated
  proof of manifest bishop finiteness to a proof of equivalence with fin.

@@

  The tricky part of the proof is \AgdaFunction{const-alg}: here we need to show
  that \AgdaFunction{alg} returns the same value no matter its input.
  That output is a pair, the first component of which is the cardinality, and
  the second the truncated equivalence proof.
  The truncated proofs in the output are trivially equal by the truncation, so
  our obligation now has been reduced to:
  \begin{equation}
    \frac{(n :\;\AgdaDatatype{\ensuremath{\mathbb{N}}}) \; \; \; (p : \AgdaDatatype{Fin}\;n\;\AgdaFunction{\ensuremath{\simeq}}\;A) \; \; \;
      (m :\;\AgdaDatatype{\ensuremath{\mathbb{N}}}) \; \; \; (q : \AgdaDatatype{Fin}\;m\;\AgdaFunction{\ensuremath{\simeq}}\;A)
    }{
      \agdaequiv{n}{m}
    }
  \end{equation}
  Given univalence we have \(\AgdaDatatype{Fin}\;n \;\AgdaFunction{\ensuremath{\equiv}}\; \AgdaDatatype{Fin}\;m\),
  and the rest of our task is to prove:
  \begin{equation}
    \frac{\agdaequiv{\AgdaDatatype{Fin}\;n}{\AgdaDatatype{Fin}\;m}}{\agdaequiv{n}{m}}
  \end{equation}

  This is a well-known puzzle in dependently-typed programming, and one that
  has a surprisingly tricky and complex proof.
  We do not include it here, since it has already been explored elsewhere, but
  it is present in our formalisation.
\end{proof}

\subsection{Going from Cardinal Finiteness to Manifest Bishop Finiteness}
We know of course that we can convert any proof of manifest Bishop finiteness to
a proof of Cardinal finiteness: it's just the truncation function
\AgdaInductiveConstructor{\(\lvert \_ \rvert\)}.
It's the other direction which presents a difficulty:
\begin{theorem} \label{cardinal-to-manifest-bishop}
  Any cardinal finite type with a total order is Bishop finite. \todo{Figure out
    how to put a type here}
\end{theorem}
\begin{proof}
The proof for this particular theorem is quite involved in the formalisation, so
we only give its sketch here \todo{ref to formalisation}.

Our strategy will be to \emph{sort} the support list of the proof for Bishop
finiteness, and then prove that the sorting function is coherently constant,
thereby satisfying the eliminator in Equation~\ref{elim-prop-coh}.
We need to show, in other words, that sorting two support lists from proofs of
manifest Bishop finiteness on the same type with the same order always returns
the same result.
For simplicity's sake we will use insertion sort:

\begin{minipage}{\linewidth}
\begin{multicols}{2}
  \begin{agdalisting}
    \ExecuteMetaData[agda/Data/List/Sort.tex]{insert}
  \end{agdalisting} \columnbreak \vfill \null \vfill
  \begin{agdalisting}
    \ExecuteMetaData[agda/Data/List/Sort.tex]{insertion-sort}
  \end{agdalisting} \vfill \null
\end{multicols}
\end{minipage} \noindent
And we prove that \AgdaFunction{sort} produces a list which is sorted, and a
permutation of its input.
\begin{multicols}{2}
  \begin{agdalisting}
    \ExecuteMetaData[agda/Data/List/Sort.tex]{sort-sorts}
  \end{agdalisting} \columnbreak
  \begin{agdalisting}
    \ExecuteMetaData[agda/Data/List/Sort.tex]{sort-perm}
  \end{agdalisting}
\end{multicols} \vspace{-\baselineskip}
We've introduced two new types here: \AgdaFunction{Sorted} is a predicate
enforcing that the given list is sorted, and
\AgdaFunction{\(\leftrightsquigarrow\)} is a permutation relation between two
lists.
We take the definition of permutations from
\cite{danielssonBagEquivalenceProofRelevant2012}: two lists are permutations of
each other if their membership proofs are all equivalent.
\begin{agdalisting}
  \ExecuteMetaData[agda/Data/List/Relation/Binary/Permutation.tex]{perm-def}
\end{agdalisting}
This definition fits particularly well for two reasons: first, it is defined on
containers generically, which fits well with our finiteness predicates.
Secondly, it is extremely straightforward to show that the support lists of any
two proofs of manifest Bishop finiteness must be permutations of each other:
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{perm-bish}
\end{agdalisting}

Almost all of the pieces are in place now: we know that the support lists of all
proofs of \(\agdacal{B}\;A\) are permutations of each other, and we know that
\AgdaFunction{sort} returns a sorted permutation of its input.
The final piece of the puzzle is the following:
\begin{agdalisting}
  \ExecuteMetaData[agda/Data/List/Sort.tex]{sorted-perm-eq}
\end{agdalisting}
If two sorted lists are both permutations of each other they must be equal.
Connecting up all the pieces we get the following:
\begin{agdalisting}
  \ExecuteMetaData[agda/Data/List/Sort.tex]{perm-invar}
\end{agdalisting}
Because we know that all support lists of \(\agdacal{B}\;A\) are permutations of
each other this is enough to prove that \AgdaFunction{sort} is coherently
constant, and therefore can eliminate from within a truncation.
The second component of the output pair (the cover proof) follows quite
naturally from the definition of permutations.
\end{proof}
\subsection{Restrictiveness}
So far our explorations into finiteness predicates have pushed us in the
direction of ``less informative'': however, as mentioned in the introduction, we
can \emph{also} ask how \emph{restrictive} certain predicates are.
Since split enumerability and manifest Bishop finiteness imply each other we
know that there can be no type which satisfies one but not the other.
We also know that manifest Bishop finiteness implies cardinal finiteness, but we
do \emph{not} have a function in the other direction:
\begin{equation} \label{c-to-b}
  \agdacal{C}\;A\rightarrow\agdacal{B}\;A
\end{equation} \todo{Solve the mystery of the mathcal difference for C and B}
So the question arises naturally: is there a cardinally finite type which is
\emph{not} manifest Bishop finite?

It turns out the answer is no!
The proof of this fact is relatively short:
\begin{agdalisting} \label{no-gap-card-bishop}
  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{no-gap-card-bishop}
\end{agdalisting}
We can apply the function of type \(\agdacal{B}\;A\rightarrow\agdabot\)
(i.e. \(\AgdaFunction{\(\neg\)}\;\agdacal{B}\;A\)) to the value of type
\(\AgdaDatatype{\(\lVert\)}\;\agdacal{B}\;A\;\AgdaDatatype{\(\rVert\)}\) (i.e.
\(\agdacal{C}\;A\)) using Equation~\ref{elim-prop}, since \agdabot\;is itself a
proposition.
This tells us that manifest bishop finiteness, cardinal finiteness, and split
enumerability all refer to the same class of types.

Interestingly, while we cannot construct a function with the type in
Equation~\ref{c-to-b}, it does exist \emph{classically}.
In fact we can derive it from Equation~\ref{no-gap-card-bishop} using
straightforward applications of De Morgan's laws:
\begin{align*}
   &\; \neg (\agdacal{C}\;A \times \neg\; \agdacal{B}\;A) \\
  =&\; \neg\; \agdacal{C}\;A + \neg \neg\; \agdacal{B}\;A \\
  =&\; \neg\; \agdacal{C}\;A + \agdacal{B}\;A \\
  =&\; \agdacal{C}\;A \rightarrow \agdacal{B}\;A
\end{align*}
\section{Manifest Enumerability} \label{manifest-enumerability}
Given that we have just proven that all of our finiteness predicates apply to
the same types, the natural next step is to try find a predicate which applies
to a different class of types.
Let's first talk about what this new class of types might look like:
what we're looking for is a type which is in some sense finite, but doesn't
conform to any of the predicates we've seen so far.
The \emph{circle} is such a type:
\begin{agdalisting}\label{circle-def}
  \ExecuteMetaData[agda/Snippets/Circle.tex]{circle-def}
\end{agdalisting}
The thing that this type has which precludes it from being, say, split
enumerable, is its \emph{higher homotopy structure}. \todo{is this the right phrase?}

@@

But the type certainly \emph{seems} finite!
It has finitely many points, for instance.
In order to explore the ``restrictiveness'' axis in
Figure~\ref{finite-classification}, then, we'll need to construct a predicate
which admits the circle.
Manifest enumerability is one such predicate.
\begin{definition}[Manifest Enumerability]
  Manifest enumerability is an enumeration predicate like Bishop finiteness or
  split enumerability with the only difference being a propositionally truncated
  membership proof.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable/Container.tex]{manifest-enum-def}
  \end{agdalisting}
\end{definition}

@@

And indeed this means that the circle is manifestly enumerable.
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{circle-is-manifest-enum}
\end{agdalisting}
We use a lemma here, proven in the Cubical Agda library, that
\AgdaDatatype{S\(^1\)} is \emph{connected}:
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{s1-connected}
\end{agdalisting}
\subsection{Surjections}
We already saw that split enumerability was the listed form of a split
surjection: what we didn't explain was why the word ``split'' was placed before
surjection.
In the presence of higher homotopies than sets, split surjections are actually
\emph{not} a satisfactory definition of surjection.
And we are most certainly in the presence of higher homotopies: just moments ago
we were introduced to the circle.
In these cases, the following definition of surjections is
preferred~\cite[definition 4.6.1]{hottbook}:
\begin{multicols}{2}
  \begin{agdalisting} \label{surj-eqn}
    \ExecuteMetaData[agda/Function/Surjective/Base.tex]{surjective}
  \end{agdalisting} \columnbreak
  \begin{agdalisting} \label{surj-arrow-eqn}
    \ExecuteMetaData[agda/Function/Surjective/Base.tex]{surjection}
  \end{agdalisting}
\end{multicols}
